From clusters to queries: exploiting uncertainty in the modularity landscape of complex networks

### 2018-06-27 article/article.tex
updates to article

% Copyright
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\acmDOI{-}

@@

%Conference
\acmConference[KDD MLG 2018]{14th International Workshop on Mining and Learning with Graphs}{Aug 20, 2018}{London, United Kingdom}
\acmYear{2018}
\copyrightyear{2018 Copyright held by the owner/author(s).}
\acmPrice{}

@@

More formally, the problem tackled in this paper can be formulated as follows:
Given a graph made up of vertices and edges $G = (V, E)$, and a query set $S \subset V$, we formally ask the question; \textit{How well is a given vertex, $i \notin S$ related to $S$?}

@@

\begin{figure}[t]
    \includegraphics[width=0.48\textwidth]{images/e_coli_search_space.eps}
    \caption{Modularity search space of an \textit{E. coli} metabolic network.
 Distance between partitions is calculated using variation of information \cite{meilua2003comparing} and dimensionality reduction is performed using curvilinear component analysis \cite{demartines1997curvilinear}.
  The inset (top) demonstrates the landscape of the high modularity region.
  Figure generated with the software of Good \textit{et al.} \cite{good2010performance}.}
    \label{fig:modular_search_space}
\end{figure}

@@

\subsection{Random walk with restart}
In the following sections compare our method with the commonly used random walk with restart method as described in \cite{kohler2008walking}.
This simulates an infinite random walk with a fixed probability, $\alpha$ that a walker would teleport back to the initial seed nodes. 
Formally, the random walk with restart algorithm uses the form
\begin{equation} \label{eq:rwr}
 \vec{p}_{t+1} = (1 - \alpha) W \vec{p}_t + \alpha \vec{p}_0,
\end{equation}
where $W$ is the row normalised adjacency matrix of a graph, $\vec{p}_0$ is the initial walk vector such that $\vec{p}_0 = \frac{1}{|S|}$ for vertices in the seed set.
The random walk algorithm repeats equation \ref{eq:rwr} from $\vec{p}_0$ until the L1 norm between  $\vec{p}_t$ and  $\vec{p}_{t+1}$ converges to 0, simulating the steady state vector $\vec{p}_{\infty}$.
For the tests performed in this work we use a restart probability $\alpha = 0.25$.

@@

\subsubsection{Non-overlapping modules.}
Figure \ref{fig:auc_no_overlap} represents tests on 100 networks with 1000 and 5000 nodes varying the mixing coefficient (number of edges between communities).
As one would expect, the prediction of the method drops off steeply where communities are less defined above a mixing coefficient of 0.6.
Using larger seed sets also improves accuracy with some prediction of true communities being possible at extremely high levels of mixing.
Overall, results are comparable to the random walk with restart method, with slightly improved performance with a smaller number of seed nodes.

\begin{figure*}[t]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width=\textwidth]{images/lfr_binary_mo_overlap_auc_1000.eps}
        \caption{1000 nodes}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width=\textwidth]{images/lfr_binary_mo_overlap_auc_5000.eps}
        \caption{5000 nodes}
    \end{subfigure}
    \caption{Non-overlapping community varying seed nodes on LFR networks with 1000 and 5000 nodes.
     Data points represent mean AUC scores for all communities on 100 sampled networks at varying mixing coefficients.
     Error bars represent one standard deviation.}
     \label{fig:auc_no_overlap}
\end{figure*}

@@

\begin{figure*}[t]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width=\textwidth]{images/lfr_binary_overlap_auc_1000.eps}
        \caption{1000 nodes}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width=\textwidth]{images/lfr_binary_overlap_auc_5000.eps}
        \caption{5000 nodes}
    \end{subfigure}
    \caption{Overlapping community varying seed nodes on LFR networks with 1000 and 5000 nodes.
     Data points represent mean AUC scores for all communities on 100 sampled networks at varying overlapping fraction of nodes.
     Error bars represent one standard deviation.}
     \label{fig:auc_overlap}
\end{figure*}

@@

\section{Related work}
\label{sec:related_work}
This work relates very strongly to the idea of local community detection, more specifically the idea of \textit{seed set expansion} \cite{gleich2012vertex}.
Here, a given seed set is created and random walks are analysed to find clearly related communities of vertices.
One of the most common approaches to finding related vertices in a network is the random walk with restart (RWR) \cite{can2005analysis, kohler2008walking}.
This approach has been applied in fields as diverse as recommender systems and the detection of potential drug targets \cite{chen2012drug}.
In RWR the relatedness of any pair of vertices can be seen as the probability of a particle traversing the graph starting at a given vertex and ending at another.
The restart aspect of the random walk can be thought of as the probability of the walker teleporting back to the start vertex set, as a fixed parameter. 
The higher the probability of the walker ending up at a given vertex, the more likely it is that the two vertices are related.

Conceptually, RWR is very similar to the method presented in this paper given that the user has a given query.
The RWR probability as analogous to the value of $\mu_i(S)$ presented above.

The reader may also consider the similarity of this approach to that of label propagation \cite{raghavan2007near, gregory2010finding}, which seeks to find communities based on a vote, where at different time steps a node updates it's label to to be the most common amongst its neighbours.
A common problem with label propagation schemes is that they often fail to converge, resulting in many competing clusterings of a network.
We liken this to the problem encountered by Good Good \textit{et al.} \cite{good2010performance} in modularity maximisation.
The core contribution of this paper is that, in such situations, there is no single, context independent labelling scheme that can be seen as the ``true'' community structure, overlapping or not.
Indeed, the approach applied here is extremely general and could be adapted to the label propagation approach (or any other community detection algorithm) should sufficient semi-supervised group memberships be known \textit{a priori}.

>>>>

% Copyright
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\makeatletter
\renewcommand\@formatdoi[1]{\ignorespaces}
\makeatother

@@

%Conference
\acmConference[KDD MLG 2018]{14th International Workshop on Mining and Learning with Graphs}{Aug 20, 2018}{London, United Kingdom}
\acmYear{2018}
\copyrightyear{2018}
\acmPrice{}

\settopmatter{printacmref=false} % Removes citation information below abstract
\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column
\pagestyle{plain} % removes running headers

@@

More formally, the problem tackled in this paper can be formulated as follows:
Given a graph made up of vertices and edges $G = (V, E)$, and a query set $S \subset V$, we ask the question; \textit{How well is a given vertex, $i \notin S$ related to $S$?}

@@

\begin{figure}[t]
    \includegraphics[width=0.48\textwidth]{images/e_coli_search_space.eps}
    \caption{Modularity search space of an \textit{E. coli} metabolic network.
 Distance between partitions is calculated using variation of information \cite{meilua2003comparing} and dimensionality reduction is performed using curvilinear component analysis \cite{demartines1997curvilinear}.
  The inset (top) demonstrates the landscape of the high modularity partitions.
  Figure generated with the software of Good \textit{et al.} \cite{good2010performance}.}
    \label{fig:modular_search_space}
\end{figure}

@@

\subsection{Random walk with restart}
\label{sec:rwr}
In the following sections compare our method with the commonly used random walk with restart method as described in \cite{kohler2008walking}.
This simulates an infinite random walk with a fixed probability, $\alpha$ that a walker would teleport back to the initial seed nodes. 
Formally, the random walk with restart algorithm uses the form
\begin{equation} \label{eq:rwr}
 \vec{p}_{t+1} = (1 - \alpha) W \vec{p}_t + \alpha \vec{p}_0,
\end{equation}
where $W$ is the row normalised adjacency matrix of a graph, $\vec{p}_0$ is the initial walk vector such that $\vec{p}_0 = \frac{1}{|S|}$ for vertices in the seed set.
The random walk algorithm repeats equation \ref{eq:rwr} from $\vec{p}_0$ until the L1 norm between  $\vec{p}_t$ and  $\vec{p}_{t+1}$ converges to 0, simulating the steady state vector $\vec{p}_{\infty}$.
For the tests performed in this work we use a restart probability $\alpha = 0.25$.

@@

\subsubsection{Non-overlapping modules.}
Figure \ref{fig:auc_no_overlap} represents tests on 10 networks with 1000 and 5000 nodes varying the mixing coefficient (number of edges between communities).
As one would expect, the prediction of the method drops off steeply where communities are less defined above a mixing coefficient of 0.6.
Using larger seed sets also improves accuracy with some prediction of true communities being possible at extremely high levels of mixing.
Overall, results are comparable to the random walk with restart method, with slightly improved performance with a smaller number of seed nodes.

\begin{figure*}[t]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width=\textwidth]{images/lfr_binary_mo_overlap_auc_1000.eps}
        \caption{1000 nodes}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width=\textwidth]{images/lfr_binary_mo_overlap_auc_5000.eps}
        \caption{5000 nodes}
    \end{subfigure}
    \caption{Non-overlapping community varying seed nodes on LFR networks with 1000 and 5000 nodes.
     Data points represent mean AUC scores for all communities on 10 sampled networks at varying mixing coefficients.
     Error bars represent one standard deviation.}
     \label{fig:auc_no_overlap}
\end{figure*}

@@

\begin{figure*}[t]
    \centering
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width=\textwidth]{images/lfr_binary_overlap_auc_1000.eps}
        \caption{1000 nodes}
    \end{subfigure}
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \includegraphics[width=\textwidth]{images/lfr_binary_overlap_auc_5000.eps}
        \caption{5000 nodes}
    \end{subfigure}
    \caption{Overlapping community varying seed nodes on LFR networks with 1000 and 5000 nodes.
     Data points represent mean AUC scores for all communities on 10 sampled networks at varying overlapping fraction of nodes.
     Error bars represent one standard deviation.}
     \label{fig:auc_overlap}
\end{figure*}

@@

\section{Related work}
\label{sec:related_work}
This work relates very strongly to the idea of local community detection, more specifically the idea of \textit{seed set expansion} \cite{gleich2012vertex}.
Here, a given seed set is created and random walks are analysed to find clearly related communities of vertices.
One of the most common approaches to finding related vertices in a network is the random walk with restart (RWR) \cite{can2005analysis, kohler2008walking} explained in section \ref{sec:rwr}.
This approach has been applied in fields as diverse as recommender systems and the detection of potential drug targets \cite{chen2012drug}.
In RWR the relatedness of any pair of vertices can be seen as the probability of a particle traversing the graph starting at a given vertex and ending at another.
Conceptually, RWR is very similar to the method presented in this paper given that the user has a given query.
The RWR probability as analogous though not equivalent to the value of $\mu_i(S)$.

The reader may also consider the similarity of this approach to that of label propagation \cite{raghavan2007near, gregory2010finding}, which seeks to find communities based on a vote, where at different time steps a node updates it's label to to be the most common amongst its neighbours.
A common problem with label propagation schemes is that they often fail to converge, resulting in many competing clusterings of a network.
We liken this to the problem encountered by Good \textit{et al.} \cite{good2010performance} in modularity maximisation.
A core contribution of this paper is that, in such situations, there is no single, context independent labelling scheme that can be seen as the ``true'' community structure, overlapping or not.
Indeed, the approach applied here is extremely general and could be adapted to the label propagation approach (or any other community detection algorithm) should sufficient semi-supervised group memberships be known \textit{a priori}.
